% ══════════════════════════════════════════════════════════════════════════════
%  CHAPTER 1: BASICS CONCEPTS IN ASTROPHYSICS
% ══════════════════════════════════════════════════════════════════════════════
\chapter{Basic Concepts in Astrophysics}


\section{Observational Foundations}

\subsection{Specific Intensity, Flux, and Luminosity}
\begin{intuition}
Observed starlight is measured in layers: direction-resolved intensity, surface flux, and finally total emitted power.
Keeping these definitions separate avoids unit mistakes in later derivations.
\end{intuition}

\begin{theory}
\begin{itemize}
    \item Specific intensity $I_\nu$:
    \begin{equation}
        dE = I_\nu \cos\theta \, dA \, dt \, d\nu \, d\Omega
    \end{equation}
    \item Monochromatic flux density:
    \begin{equation}
        F_\nu = \int I_\nu \cos\theta \, d\Omega
    \end{equation}
    \item Bolometric flux:
    \begin{equation}
        F = \int_0^\infty F_\nu \, d\nu
    \end{equation}
    \item Luminosity:
    \begin{equation}
        L = \int F \, dA
    \end{equation}
\end{itemize}
\end{theory}

\begin{importantnote}
$I_\nu$ is invariant along a ray in vacuum (no absorption/emission), while $F_\nu$ depends on source geometry and distance.
Do not interchange them in magnitude or distance calculations.
\end{importantnote}

\subsection{Magnitude System and Distance Modulus}
\begin{theory}
\begin{align}
    M_1 - M_2 &= -2.5 \log_{10}\left(\frac{L_1}{L_2}\right), \\
    m_1 - m_2 &= -2.5 \log_{10}\left(\frac{F_1}{F_2}\right), \\
    m - M &= 5 \log_{10}(d/\text{pc}) - 5, \\
    F &= \frac{L}{4\pi d^2}.
\end{align}
\end{theory}

\begin{example}{Distance modulus quick use}{ex:distmod}
A star has $m=9.0$ and $M=4.0$. Estimate its distance.

\smallskip\textbf{Solution.}
\begin{align*}
m-M &= 5\log_{10}(d/\text{pc})-5 = 5 \\
\Rightarrow \log_{10}(d/\text{pc}) &= 2 \\
\Rightarrow d &= 10^2\,\text{pc} = 100\,\text{pc}.
\end{align*}
\end{example}


\subsection{Parallax and Radius}
\begin{theory}
\begin{itemize}
    \item Parallax distance:
    \begin{equation}
        d(\text{pc}) = \frac{1}{\pi(\text{arcsec})}
    \end{equation}
    \item Radius from angular diameter $\theta$:
    \begin{equation}
        R = \frac{1}{2}\theta d
    \end{equation}
\end{itemize}
\end{theory}
\begin{example}{}{ex:distmod2}
From the previous example, if the star has an angular diameter of $0.01$ arcseconds, estimate its radius in units of $R_\odot$.

\smallskip\textbf{Solution.}
\begin{align*}
R &= \frac{1}{2}\theta d = \frac{1}{2} \times 0.01'' \times 100\,\text{pc} \\
&= 0.5 \times 0.01 \times 100 \times 206265\,\text{AU} \\
&= 1031.325\,\text{AU} \\   
&= 1031.325 \times 215 R_\odot \\
&\approx 221,000 R_\odot.
\end{align*}
\end{example}



\section{Hydrostatic Equilibrium}
\begin{foundation}
Assume spherical symmetry and radial dependence only.
Neglect rotation, magnetic fields, and turbulence in the first model to isolate gravity-versus-pressure physics.
\end{foundation}

% ═════════════════════════════════════════════════════════════════════════════ 
\section{Stars and Stellar Properties}
\begin{foundation}
    In this notes, we define a star as a celestial object that is:
    \begin{itemize}
        \item (1) bounded by self-gravity under hydrostatic equilibrium, and
        \item (2) radiating energy generated by internal sources (e.g. nuclear fusion).
    \end{itemize}
    For most of the case, a star is spherical due to spherical symmetry of gravity, and mainly radiating energy by nuclear fusion in the core. However, we will also encounter some exceptions (e.g. energy generated by gravitational contraction in late stages of star evolution). \\

    This definition shows that other celestial objects (e.g. planets, comets) are not stars as they don't fulfill all conditions. On the other hand, it allow some special conditions (e.g. brown dwarfs, white dwarfs) to be classified as stars, as they are still self-gravitating and radiating energy (though not by nuclear fusion). However, be careful some astrophysicists may not treat them as "real stars" as they don't have nuclear fusion, but we will still include them in our discussion as they are important for understanding the full picture of stellar evolution.
\end{foundation}

% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Distance and Parallax}


% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Brightness, Magnitude, and Luminosity}


% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Mass}







% ═════════════════════════════════════════════════════════════════════════════ 
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\section{Thermal Radiation and Blackbody Spectrum}
% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Planck function, Rayleigh-Jeans law, Wien's law}


% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Intensity, Flux, and Luminosity}


% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Radiative Transfer}


% ---------------------------------------------------------------------------------
\bigskip\noindent \bigskip\noindent \bigskip\noindent
\subsection{Surface Temperature}







