\chapter{Thermonuclear Fusion in Stars}

Recall that there are three primary energy sources in stars: gravitational contraction, radioactive decay and nuclear fusion. 

In this chapter, we will focus on the process of thermonuclear fusion, which is the viable long-term stellar energy source in most stars.

\section{Nuclear Basics}

\subsection{Nuclear Size and Binding Energy}
\begin{theory}
\begin{align}
    R &= R_0A^{1/3}, \qquad R_0\approx1.2\times10^{-15}\,\text{m}, \\
    BE(Z,N) &= \left[Zm_p+Nm_n-M(Z,N)\right]c^2.
\end{align}
Binding energy per nucleon increases for light nuclei and decreases for very heavy nuclei due to Coulomb repulsion.
\end{theory}

\begin{example}{}{ex:binding}
Calculate the binding energy per nucleon for helium-4 ($^4$He) and $^{12}$C.

\smallskip\textbf{Solution.}
\begin{align*} 
BE(^4\text{He}) &= \left[2m_p + 2m_n - M(^4\text{He})\right]c^2 \approx 28.3\,\text{MeV}, \\
BE(^4\text{He})/4 &\approx 7.1\,\text{MeV/nucleon}, \\
BE(^{12}\text{C}) &= \left[6m_p + 6 m_n - M(^{12}\text{C})\right]c^2 \approx 92.2\,\text{MeV}, \\
BE(^{12}\text{C})/12 &\approx 7.7\,\text{MeV/nucleon}.
\end{align*}   
\end{example}

\subsection{Why Fusion Powers Stars}
\begin{intuition}
Stars gain energy when fusing light nuclei toward higher binding energy per nucleon.
This trend reverses near iron, where further fusion is endothermic.
\end{intuition}

\section{Hydrogen and Helium Burning}
\subsection{Proton-Proton (p-p) Chain}

\begin{foundation}
Representative net reaction:
\begin{equation}
    4\,{}^1\text{H}\rightarrow{}^4\text{He}+2e^++2\nu_e+\text{energy}.
\end{equation}
Total released energy is about $26.7\,$MeV, with part carried away by neutrinos.
\end{foundation}

\section{CNO Cycle}


\section{Barrier Penetration and Reaction Rates}

\subsection{Coulomb Barrier and Tunneling}
Derivation

\begin{foundation}
The \textbf{Gamov energy} is the minimum kinetic energy that two nuclei must have in order to overcome the Coulomb barrier and undergo fusion. It is given by the formula:
\begin{equation}    
E_G = (\pi \alpha Z_1 Z_2)^2 2 m_{r}c^2
\end{equation} 
where $m_r$ is the reduced mass of the two nuclei, $\alpha$ is the fine-structure constant, and $Z_1$ and $Z_2$ are the atomic numbers of the two nuclei.
\end{foundation}
\begin{example}{The fusion of two protons}{ex:gamov1}
Calculate the Gamov energy $E_G$ for the fusion of two protons.

\smallskip\textbf{Solution.}
For two protons, we have $Z_1 = Z_2 = 1$ and the reduced mass $m_r$ is approximately equal to the mass of a proton, $m_p$. The fine-structure constant $\alpha$ is approximately $1/137$. Plugging these values into the formula for the Gamov energy, we get:
\begin{align*}
E_G &= (\pi \alpha Z_1 Z_2)^2 2 m_{r}c^2 \\
&= (\pi \times \frac{1}{137} \times 1 \times 1)^2 \times 2 m_p c^2 \\
&= \left(\frac{\pi}{137}\right)^2 \times 2 m_p c^2 \\
&\approx 493\,\text{keV}.
\end{align*}
Thus, the Gamov energy for the fusion of two protons is approximately $493\,\text{keV}$.
\end{example}

\begin{example}{The fusion of carbon and oxygen}{ex:gamov2}
Calculate the Gamov energy $E_G$ for the fusion of carbon-12 ($^{12}\text{C}$) and oxygen-16 ($^{16}\text{O}$). 

\smallskip\textbf{Solution.}
For carbon-12, we have $Z_1 = 6$ and for oxygen-16, we have $Z_2 = 8$. The reduced mass $m_r$ can be calculated using the formula:
\begin{equation}
m_r = \frac{m_1 m_2}{m_1 + m_2},
\end{equation}
where $m_1$ and $m_2$ are the masses of the carbon and oxygen nuclei, respectively. The mass of carbon-12 is approximately $12 m_p$ and the mass of oxygen-16 is approximately $16 m_p$. Thus, we have:
\begin{align*}
m_r &= \frac{(12 m_p)(16 m_p)}{12 m_p + 16 m_p} \\
&= \frac{192 m_p^2}{28 m_p} \\        
&= \frac{192}{28} m_p \\
&\approx 6.86 m_p.
\end{align*}
Now we can calculate the Gamov energy using the formula:
\begin{align*}
E_G &= (\pi \alpha Z_1 Z_2)^2 2 m_{r}c^2 \\
&= (\pi \times \frac{1}{137} \times 6 \times 8)^2 \times 2 \times 6.86 m_p c^2 \\
&= \left(\frac{48\pi}{137}\right)^2 \times 2 \times 6.86 m_p c^2 \\
&\approx 8.7\,\text{MeV}. 
\end{align*}
Thus, the Gamov energy for the fusion of carbon-12 and oxygen-16 is approximately $8.7\,\text{MeV}$.
\end{example}


\begin{foundation}
Classically, nuclei with $E<V_C$ cannot fuse.
Quantum tunneling gives non-zero penetration probability $P$ for $E<V_C$. The probability of tunneling through the Coulomb barrier can be approximated by the formula:
\begin{equation}
    P \propto \exp\left[-\sqrt{\frac{E_G}{E}}\right],
\end{equation}
where $E_G$ is the Gamow energy.
\end{foundation}

\section{Thermonuclear Reaction Rates}

\begin{theory}
\begin{equation}
    \langle\sigma v\rangle = \left(\frac{8}{\pi\mu}\right)^{1/2}
    \frac{1}{(k_BT)^{3/2}}\int_0^\infty S(E)
    \exp\left[-\frac{E}{k_BT}-\sqrt{\frac{E_G}{E}}\right]dE.
\end{equation}
The integrand peaks at the Gamow energy window, balancing high-energy tail abundance and tunneling probability.
\end{theory}


\begin{theory}
\begin{equation}
    R_{AB} \propto T^{a}
\end{equation}
where $a = \sqrt[3]{\frac{E_G}{4kT}}$ is a positive exponent that depends on the specific reaction and the temperature range.
\end{theory}

\begin{example}{}{ex:ppchain}
Calculate the temperature dependence of the proton-proton (p-p) chain reaction rate.

\smallskip\textbf{Solution.}
The proton-proton chain reaction has a Gamow energy of approximately $E_G \approx 493\,\text{keV}$. The temperature dependence can be approximated by:
\begin{equation}
    R_{pp} \propto T^{\sqrt[3]{\frac{E_G}{4kT}}}.           
\end{equation}
At typical stellar core temperatures of around $T \approx 1.5 \times 10^7\,\text{K}$, the exponent $a$ can be calculated as:
\begin{equation}
    a = \sqrt[3]{\frac{493\,\text{keV}}{4k(1.5 \times 10^7\,\text{K})}} \approx 4.5.    
\end{equation}
Thus, the reaction rate for the proton-proton chain scales approximately as $R_{pp} \propto T^{4.5}$ at typical stellar core temperatures.
\end{example}