% ══════════════════════════════════════════════════════════════════════════════
%  CHAPTER 5: NUCLEAR SYNTHESIS IN STARS
% ══════════════════════════════════════════════════════════════════════════════
\chapter{Nuclear Synthesis in Stars}

% ═════════════════════════════════════════════════════════════════════════════ 

Recall that there are three primary energy sources in stars: gravitational contraction, radioactive decay and nuclear fusion. 

In this chapter, we will focus on the process of thermonuclear fusion, which is the viable long-term stellar energy source in most stars.

\section{Nuclear Basics}

\subsection{Nuclear Size and Binding Energy}
\begin{theory}
\begin{align}
    R &= R_0A^{1/3}, \qquad R_0\approx1.2\times10^{-15}\,\text{m}, \\
    BE(Z,N) &= \left[Zm_p+Nm_n-M(Z,N)\right]c^2.
\end{align}
Binding energy per nucleon increases for light nuclei and decreases for very heavy nuclei due to Coulomb repulsion.
\end{theory}

\subsection{Why Fusion Powers Stars}
\begin{intuition}
Stars gain energy when fusing light nuclei toward higher binding energy per nucleon.
This trend reverses near iron, where further fusion is endothermic.
\end{intuition}

\section{Hydrogen and Helium Burning}
\subsection{Proton-Proton (pp) Chain}

\begin{foundation}
Representative net reaction:
\begin{equation}
    4\,{}^1\text{H}\rightarrow{}^4\text{He}+2e^++2\nu_e+\text{energy}.
\end{equation}
Total released energy is about $26.7\,$MeV, with part carried away by neutrinos.
\end{foundation}

\section{CNO Cycle}


\section{Barrier Penetration and Reaction Rates}

\subsection{Coulomb Barrier and Tunneling}
Derivation

\begin{foundation}
The \textbf{Gamov energy} is the minimum kinetic energy that two nuclei must have in order to overcome the Coulomb barrier and undergo fusion. It is given by the formula:
\begin{equation}    
E_G = (\pi \alpha Z_1 Z_2)^2 2 m_{r}c^2
\end{equation} 
where $m_r$ is the reduced mass of the two nuclei, $\alpha$ is the fine-structure constant, and $Z_1$ and $Z_2$ are the atomic numbers of the two nuclei.
\end{foundation}

\begin{foundation}
Classically, nuclei with $E<V_C$ cannot fuse.
Quantum tunneling gives non-zero penetration probability $P$ for $E<V_C$. The probability of tunneling through the Coulomb barrier can be approximated by the formula:
\begin{equation}
    P \propto \exp\left[-\sqrt{\frac{E_G}{E}}\right],
\end{equation}
where $E_G$ is the Gamow energy.
\end{foundation}

\section{Thermonuclear Reaction Rates}

\begin{theory}
\begin{equation}
    \langle\sigma v\rangle = \left(\frac{8}{\pi\mu}\right)^{1/2}
    \frac{1}{(k_BT)^{3/2}}\int_0^\infty S(E)
    \exp\left[-\frac{E}{k_BT}-\sqrt{\frac{E_G}{E}}\right]dE.
\end{equation}
The integrand peaks at the Gamow energy window, balancing high-energy tail abundance and tunneling probability.
\end{theory}


\begin{theory}
\begin{equation}
    R_{AB} \propto T^{a}
\end{equation}
where $a = \sqrt[3]{\frac{E_G}{4kT}}$ is a positive exponent that depends on the specific reaction and the temperature range.
\end{theory}